---
title: "Supplementary information to *Satellite observations underestimate the impact of drought on terrestrial primary productivity*"
author: "Benjamin D. Stocker, Jakob Zscheischler, Trevor F. Keenan, I. Colin Prentice, Josep Penuelas, and Sonia I. Seneviratne"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: true
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
#     number_sections: true
    includes:
      in_header: header.tex
header-includes:
   - \usepackage{amsmath}
bibliography: bibliography.bib
csl: nature.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(captioner)
tab_nums <- captioner(prefix = "Table S", auto_space=FALSE, style = "i")
fig_nums <- captioner(prefix = "Figure S", auto_space=FALSE, style = "i")
```

This document provides supplementary information to the article *Satellite observations underestimate the impact of drought on terrestrial primary productivity*. References are made to scripts and functions implemented in R and available open access through github (https://github.com/stineb/soilm_global). This repository also provides a detailed workflow description as an RMarkdown file (`knit_soilm_global.Rmd`) that allows for full reproducibility of published results and figures, including contents of this document.


# Data preparation

## Data collection

First, we combined FLUXNET 2015 data (daily, from Tier 1 sites), and respective environmental forcing data (meteorological variables from FLUXNET, fAPAR from MODIS FPAR, and soil moisture distributed through the FLUXNET 2015 dataset).

We use GPP data, derived from measurements of net ecosystem CO$_2$ exchange based on the nighttime partitioning method [@Reichstein2005-mp], named `GPP_NT_VUT_REF` in the FLUXNET 2015 dataset. We filtered negative daily GPP values, data for which more than 50% of respective half-hourly data is gap-filled and for which the daytime and nighttime methods (`GPP_DT_VUT_REF` and `GPP_NT_VUT_REF`, resp.) are inconsistent, i.e., the upper and lower 2.5% quantile of the difference between each method's GPP quantification.

This step is implemented by code `get_modobs.R`.

## Data aggregation

We combine site-level observational data with site-level model outputs from the P-model, BESS, VPM, and MODIS, and aggregate data to 8-days periods, defined by MODIS data frequency.

This step is implemented by code `aggregate_nn_fluxnet2015.R`

## Data filtering for soil moisture effects

The comparison of GPP from RS-models and FLUXNET observations is limited to data from a subset of sites, where soil moisture effects had reliably been identified (36 sites), and to periods with clearly identified soil moisture effects. The site selection is determined by the amount of data available from respective sites (at least 500 days' data for the site after data cleaning) and the importance of soil moisture effects on observed LUE changes. This defines 'group 1' in Table S1 and is based on the analysis by [@Stocker2018-uf]. An overview of selected sites is given also by Figure 1. A wider set of sites with relaxed criteria (71 sites, group 2 in Table S1) is used for confirming the systematic bias in soil dryness and the bias in GPP estimates. This set includes additional sites with good quality soil moisture and vegetation greennes data available along with sufficient flux data. Sites of group 1 are displayed in `r fig_nums( "siteoverview_fig", display="cite" )` as green dots, sites of group 2 as black dots.

```{r siteoverview, echo=FALSE, warning=FALSE}
suppressMessages(require(readr, quietly = TRUE))
suppressMessages(require(dplyr, quietly = TRUE))
siteinfo <- suppressMessages(read_csv("siteinfo_fluxnet2015_sofun.csv"))
success  <- suppressMessages(read_csv( "successcodes.csv" ))
do.sites.nn <- filter( success, successcode==1 )$mysitename
do.sites.all <- filter( success, successcode==1 | successcode==2 )$mysitename
siteinfo <- siteinfo %>%  mutate( group = ifelse(  mysitename %in% do.sites.nn, 1, ifelse(  mysitename %in% do.sites.all, 2, 0 ) ) ) %>%
              select( -elv, -elv_watch, -elv_diff, -years_data, -whc ) %>%
              dplyr::rename( site=mysitename ) %>%
              filter( group %in% c(1,2) ) %>%
              arrange( group ) %>%
              mutate( citation = paste0("[@", site ,"]") ) %>%
              mutate( Period=paste0(as.character(year_start), "-", as.character(year_end)) )
siteinfo %>%  select( -year_start, -year_end ) %>%
              dplyr::rename( Site=site, Lon.=lon, Lat.=lat, Veg.=classid, Group=group, Reference=citation ) %>%
              select( Site, Lon., Lat., Veg., Period, Group, Reference ) %>%
              knitr::kable( caption = "Sites used for bias evaluation. Lon. is longitude, negative values indicate west longitude; Lat. is latitude, positive values indicate north latitude; Veg. is vegetation type: deciduous broadleaf forest (DBF); evergreen broadleaf forest (EBF); evergreen needleleaf forest (ENF); grassland (GRA); mixed deciduous and evergreen needleleaf forest (MF); savanna ecosystem (SAV); shrub ecosystem (SHR); wetland (WET)." )
cap_siteoverview_tab <- tab_nums( "siteoverview_tab", caption=" Sites used for bias evaluation. Lon. is longitude, negative values indicate west longitude; Lat. is latitude, positive values indicate north latitude; Veg. is vegetation type: deciduous broadleaf forest (DBF); evergreen broadleaf forest (EBF); evergreen needleleaf forest (ENF); grassland (GRA); mixed deciduous and evergreen needleleaf forest (MF); savanna ecosystem (SAV); shrub ecosystem (SHR); wetland (WET)." )
```



```{r siteoverview_fig, echo=FALSE, warning=FALSE}
suppressMessages(require(ncdf4, quietly = TRUE))
ncfiln <- "../data/greve/ep_over_p_cru_ncep.nc"
if (!file.exists(ncfiln)) {
  epop <- array( 1, dim=c(720,360) )
} else {
  nc <- nc_open( ncfiln )
  epop <- ncvar_get( nc, varid="EP_OVER_P_CRU_NCEP" )
}
source("plot_map_siteoverview.R")
suppressMessages( plot_map_siteoverview( siteinfo, 1/epop ) ) # , plotfiln="fig/map_sites.pdf"
cap_siteoverview_fig <- fig_nums( "siteoverview_fig", caption=" Geographical distribution of sites selected for the bias evaluation. Sites listed in Table S1 as group 1 are in green, sites of group 2 are in black. The color of land area represents aridity, quantified as the ratio of precipitation over potential evapotranspiration  from Greve et al.[@greve14]" )
```

`r fig_nums("siteoverview_fig")`

In a second step, data is filtered for periods where soil moisture impacts are significant based on the parallel evolution of observed photosynthetic light use efficiency (LUE), temperature, vapour pressure deficit (VPD), photosynthetically active radiation (PAR), and soil moisture. Impact-relevant drought periods are defined when LUE is significantly reduced by soil moisture effects ('fLUE droughts' in [@Stocker2018-uf]). Only data is retained for the 20-day period prior to the onset of fLUE droughts and during the duration of fLUE droughts.

The data filtering steps are implemented by code `execute_reshape_align_nn_fluxnet2015.R`.

For additional analyses with relaxed criteria for site selection and independent of the drought identification by Stocker et al. (2018)[@Stocker2018-uf], alternative data aggregation is implemented by code `aggregate_all_fluxnet2015.R`.


# Bias detection

We evaluate the bias of different RS-model versus actual over potential evapotranspiration (AET/PET) for 71 sites (corresponding to the full list of sites shown in Table S1). This corresponds to the selection with relaxed criteria and is independent of the drought identification by Stocker et al. (2018)[@Stocker2018-uf]. For all RS-models, the bias generally increases along bins of decreasing AET/PET. Note that in `r fig_nums("bias_all", display="cite")`, boxplots are shown for the ratio of modelled over observed GPP.

```{r plot_bias_all, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
source("plot_bias_all.R")
plot_bias_all( nice_agg, nice_8d_agg )
cap_bias_all <- fig_nums( "bias_all", caption=" Bias of simulated daily GPP by different RS-models and bins of daily AET/PET. The bias is calculated here as the ratio of simulated GPP over observed GPP. Observational data is derived from flux measurements at 71 sites (all sites in Table S1)." )
```

`r fig_nums("bias_all")`

This pattern becomes clearer when applying the narrower site selection (sites in group 1 in Table S1) and using only data during apparent droughts, i.e. where soil-moisture appears to reduce the LUE as determined by Stocker et al. (2018)[@Stocker2018-uf].

```{r echo=FALSE, warning=FALSE, error=FALSE, plot_bias_problem}
source("plot_bias_problem.R")
plot_bias_bymodels( df_dday_8d_agg )
cap_bias_bymodels <- fig_nums("bias_bymodels", caption=" Bias of simulated daily GPP by models, in different fLUE bins, calculated as simulated GPP minus observed GPP, derived from flux measurements at 36 sites (sites of group 1 in Table S1).")
```

`r fig_nums("bias_bymodels")`

All RS-models, except P-model, have a tendency to be biased low under non-soil moisture-limited conditions and biased under strong soil moisture stress. To distill the pattern, independent of the mean absolute bias, we normalised simulated GPP to the median ratio of modelled-over-observed GPP in the highest fLUE bin, i.e., where soil moisture impacts on observed fluxes are small (fLUE bin 0.8-1.0). This is shown by Fig. 1 in the main text, where the bias values of all RS models are aggregated, and by `r fig_nums( "bias_bymodels_norm", display="cite" )` below.

```{r echo=FALSE, warning=FALSE, error=FALSE, plot_bias_problem_norm}
plot_bias_bymodels( df_dday_8d_agg_norm ) #, filn="fig/bias_bymodels_norm.pdf"
cap_bias_bymodels <- fig_nums("bias_bymodels_norm", caption=" Bias of simulated daily GPP by models, in different fLUE bins, calculated as simulated GPP minus observed GPP, derived from flux measurements at 36 sites (sites of group 1 in Table S1). Bias values are normalised to a median of 1 for the 20 days before  the onset of drought events.")
```

`r fig_nums("bias_bymodels_norm")`

# Supplementary figures and analyses.

## Site-by-site evaluations of $\beta$ functions

Site-by-site evaluations of the three different soil moisture stress functions against fLUE data from [@Stocker2018-uf] are shown below in `r fig_nums("fig_fr", display="cite")`, below.

```{r plot_site_vs_soilmoisture, fig.width=3.8, fig.height=3.5, echo=FALSE, warning=FALSE, error=FALSE}
source("plot_fit_vs_soilmoist.R")
load("data/linearfit_I.Rdata")
load("data/linearfit_IV.Rdata")
load("data/linearfit_III.Rdata")
load("data/ddf.Rdata")
load("data/nice_agg.Rdata")
plot_fit_vs_soilmoist( linearfit_I, linearfit_IV, linearfit_III, ddf, nice_agg, makepdf = FALSE )
fig_fr <- fig_nums("fig_fr", caption=" Illustration of the different empirical soil moisture stress functions (lines) and apparent soil moisture-related reductions in light use efficiency (fLUE, dots).")
```

`r fig_nums("fig_fr")`


## Global GPP trends

In all simulations, global GPP exhibits an increasing trend over the years 1982-2016 in our simulations. Simulation *s0* does not account for soil moisture effect and respective global GPP has a positive trend of 0.52 PgC yr$^{-1}$. Accounting for soil moisture stress by factoring in $\beta$ functions reduces GPP by 10-19$\%$ and reduces the linear trend in global GPP is reduced to 0.45 PgC yr$^{-1}$ in s1b. The relative reduction by soil moisture effects is stable across time. That is, it doesn't exhibit a significant trend, as illustrated by `r fig_nums( "tseries_red", display="cite")` below.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
source("plot_effects_gpp_tseries.R")
plot_gpp_global_tseries_relefftrend_pmodel( df )
cap_tseries_red <- fig_nums( "tseries_red", caption=" Reduction in global annual GPP due to soil moisture effects (in %) and linear trend in different P-model simulations. The slope is given in %/yr. The red range is given by simulations s1a and s1c, with s1b given by the red line in its center." )
```

`r fig_nums( "tseries_red") `

<!-- Soil moisture effects decrease global GPP on average by: -->
<!-- ```{r, echo=FALSE, warning=FALSE, message=FALSE} -->
<!-- source("plot_effects_gpp_tseries.R") -->
<!-- get_numbers_effects_gpp_tseries( df ) -->
<!-- ``` -->

## Spatial and interannual variability

```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, fig.width=5, fig.height=4}
## Get annual values and evaluate power of getting spatial and annual variations
suppressMessages(source("aggregate_nn_fluxnet2015_annual.R"))
plot_spatial_annual_s0( adf, meandf, linmod_list_pmodel_s0,  linmod_mean_pmodel_s0,  stats_adf, stats_meandf, slope_wgt_pmodel_s0 )
fig_sa_s0 <- fig_nums("fig_sa_s0", caption=" Modelled and observed annual GPP in the simulation ignoring direct soil moisture effects (simulation s0). Red line and text is based on means across years by site and represents spatial (across-site) variations. Black lines and text is based on annual values, one line for each site. Lines represent linear regressions. R$^2$ and RMSE statistics for annual values (black text) are based on pooled data from all sites; a histogram of regression slope for all sites is given by the inset in the top-left corner.")
```

`r fig_nums("fig_sa_s0")`

```{r echo=FALSE, warning=FALSE, error=FALSE, fig.width=5, fig.height=4}
## Get annual values and evaluate power of getting spatial and annual variations
plot_spatial_annual_s1( adf, meandf, linmod_list_pmodel_s1b,  linmod_mean_pmodel_s1b,  stats_adf, stats_meandf, slope_wgt_pmodel_s1b )
fig_sa_s1 <- fig_nums("fig_sa_s1", caption=" Modelled and observed annual GPP in the simulation ignoring accounting for soil moisture effects (method IV). Red line and text is based on means across years by site and represents spatial (across-site) variations. Black lines and text is based on annual values, one line for each site. Lines represent linear regressions. R$^2$ and RMSE statistics for annual values (black text) are based on pooled data from all sites; a histogram of regression slope for all sites is given by the inset in the top-left corner." )
```

`r fig_nums("fig_sa_s1")`

## Amplification of absolute variability

```{r, echo=FALSE, warning=FALSE, message=FALSE}
source("map_effects_gpp_var.R")

# plot_map( gpp_s1/gpp_s0, lev=lev, maxval=maxval, color = out.mycolorbar$colors )
# cap_ampl_var <- fig_nums( "ampl_var", caption=" Amplification (unitless) of GPP IAV due to the effects of soil moisture. IAV is calculated as the variance in annual values; ???amplification??? is calculated as the ratio in IAV of simulation s1b to s0." )

plot_map( gpp_s1b - gpp_s0, lev=seq(-5000,13000,2000), minval=-169000, maxval=83000, toplefttext=expression(paste("")), color=c( "royalblue4", "wheat", "tomato2", "tomato4" )  ) #, toprighttext=expression(paste("g C m"^-2, " yr"^-1 ) )
cap_ampl_var <- fig_nums( "ampl_var", caption=" Difference in GPP IAV due to the effects of soil moisture (gC m-2 yr-1). IAV is calculated as the variance in annual values. The difference is taken from simulations s0 and s1b." )
```

`r fig_nums( "ampl_var") `

## Amplification of relative variability by mean aridity

```{r, echo=FALSE, warning=FALSE, message=FALSE}
source("map_effects_gpp_relvar.R")
par(las=1)
myboxplot( ampl ~ inbin, data = df_pixels, ylim=c(0,5.5), outline=FALSE,
           col=colorRampPalette( rev( c( "royalblue4", "wheat", "tomato2", "tomato4" ) ) )( 6 ),
           ylab="Amplification", xlab="Aridity index")
abline(h=1, lty=3)
cap_ampl_var <- fig_nums( "ampl_boxpl", caption=" Amplification of relative variance in GPP due to soil moisture effects versus aridity index, calculated as the ratio of precpipitation over potential evapotranspiration from [@greve14]." )
```

`r fig_nums( "ampl_boxpl") `

## Trend in GPP extreme event impacts

```{r, echo=FALSE, warning=FALSE, message=FALSE}
load("data/df_impact_year.Rdata")
lm_slope <- function(linmod){
  ci <- confint(linmod)
  eq <- substitute( italic("slope") == a~ "["*b~ - ~c*"] yr" ^-1,
                    list(a = format(coef(linmod)[2], digits = 2), b = format( ci[2,1], digits = 2 ), c = format( ci[2,2], digits = 2 ) ) )
  as.character(as.expression(eq))
}
linmod <- lm( diff ~ year, data = df_impact_year )
ggp <- ggp <- ggplot( df_impact_year, aes( x = year, y = diff ) ) +
  geom_point() +
  geom_smooth (method = "lm", level = 0.95 ) +
  theme_bw() +
  labs( x = "Year", y = bquote( "Impact difference ("*Pg~ C~ yr^-1*")") ) +
  annotate( geom = "text", x = 1997, y = 0.7, label = lm_slope(linmod), parse = TRUE, adj = 0 )
last_plot()
cap_ampl_var <- fig_nums( "ampl_extr", caption=" Difference in the impact of the 200 largest GPP extreme events in years 1982-2016. The difference is calculated from the impact of all events per year based on simulations s0 and s1b." )
```

`r fig_nums( "ampl_extr") `


# Global P-model simulations

## Simulation setup

Empirical soil moisture stress functions $\beta_a$, $\beta_b$, and $\beta_c$ are implemented for global simulations with the P-model. Simulation *s1a* uses $\beta_a$, *s1b* uses $\beta_b$, and *s1c* uses $\beta_c$.

## Model

The P-model [@Wang2017-ls] and SPLASH [@Davis2017-jo] are implemented for site-scale and global applications within a single flexible modelling framework, SOFUN v1.0.0 [@sofun_v110] available through Zenodo(https://zenodo.org/record/1213758#.WubnudNubOQ) and Github (https://github.com/stineb/sofun/tree/v1.1.0). This is based on the formulation to predict leaf-level photosynthetis and light use efficiency of (P-model) [@Wang2017-ls] and the SPLASH model [@Davis2017-jo] for calculating potential and actual evapotranspiration and simulating the soil water balance. Different implementations of the P-model have been used previously [@Keenan2016-ax]. A monthly LUE is calculated based on monthly mean climate conditions, calculated from daily values (see below). Daily GPP is then calculated based on monthly LUE multiplied by daily varying absorbed PAR and aggregated again to longer periods for all evaluations. The time scale at which the light use efficiency concept can be applied is determined by the time scale of allocation into the photosynthetic machinery (turnover of enzymes). The linearity between primary production and absorbed light described by Monteith's light use efficiency model (Eq. 1 in main text) arises due to acclimation of photosynthesis to light conditions but does not hold at short time scales (e.g., hourly-daily), where the light response curves are strongly non-linear.

## Greenness data

The P-model is used here as a light use efficiency model [@Monteith1972-ee] (Eq. 1 in main text), driven by remotely sensed greenness (fAPAR). We use an updated version of FPAR3g [@Zhu2013-hs] here, which covers years 1982-2016 and limits model simulations into the past.

## Climate data

Daily data from the WATCH-WFDEI climate forcing dataset [@Weedon2014-nv] is used here for variables temperature, precipitation (sum of rainfall and snowfall), specific humidity (converted to vapour pressure deficit), and shortwave incoming radiation (converted to photosynthetically active radiation).

### Vapour pressure deficit
Vapour pressure deficit (VPD, here $D$, in units of Pa) is calculated from specific humidity ($q$) as:
$$
D = e_s - e_a
$$
wehere $e_s$ is the saturation water vapour pressure, and $e_a$ is the actual water vapour pressure (both in Pa). $e_s$ is calculated as:
$$
e_s = 611.0 \; \exp \Big( \frac{17.27 \; T}{T + 237.3} \Big)
$$
$T$ is temperature. Actual water vapour pressure $e_a$ is calculated as:
$$
e_a = P\frac{w R_v}{R_d + w R_v}
$$
where $P$ is the atmospheric pressure, $w$ is the mass mising ratio of water vapor to dry air (dimensionless), $R_v$ is the specific gas constant of water vapor (J g$^{-1}$ K$^{-1}$), and $R_d$ is the specific gas constant of dry air (J g$^{-1}$ K$^{-1}$). $P$ is calculated from standard conditions and spatially varying elevation. $w$ is calculated from specific humidity $q$ data:
$$
w = q/(1-q).
$$


### Photosynthetically active radiation

Photosynthetically active radiation (PAR, in units of mol m$^{-2}$ d$^{-1}$) is calculated from shortwave incoming radiation $R_s$ (W m$^{-2}$) as
$$
\text{PAR} = R_s \; k \; 60.0 \cdot 60.0 \cdot 24.0 \cdot  10^{-6}
$$
where $k$ is a flux-to-energy conversion factor, here 2.04 $\mu$mol J$^{-1}$.

## Soil data

Plant-available soil water holding capacity (WHC) is an essential variable that determines water stress during dry periods. To estimate variations in WHC across the globe, we used texture data from SoilGrids [@Hengl2014-jm] at 1 km resolution and derived plant wilting point $F_{\text{PWP}}$ and field capacity $F_{\text{FC}}$ estimated from sand, clay and organic matter contents following [@Saxton2006-zb]. WHC is calculated as
$$
\text{WHC} = (F_{\text{PWP}}-F_{\text{FC}})(1-f_{\text{gravel}})\cdot\min(z, z_{\text{max}})
$$
where $f_{\text{gravel}}$ is the coarse fraction (SoilGrids data) and $z$ is the soil depth to bedrock (SoilGrids data) and $z_{\text{max}}$ is the maximum soil depth considered here, taken as 2000 mm. For global applications, global WHC fields derived at 1 km resolution are aggregated to 0.5$^{\circ}$ in longitude and latitude.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
filn <- paste0( myhome, "data/soil/soilgrids/KWm_halfdeg.nc")
if (file.exists(filn)){
  suppressMessages(library(ncdf4))
  suppressMessages(library(pracma))
  source("plot_map.R")
  nc <- nc_open( filn )
  whc <- ncvar_get( nc, varid="KWm" )
  whc <- flipdim( whc, 2 )
  nc_close( nc )
  plot_map( whc, lev=c(0,400,10), color=c( "wheat", "royalblue4" ))
  fig_whc <- fig_nums("fig_whc", caption=" Water holding capacity (mm) used for P-model simulations.")
} else {
  print("WHC file not available. Contact the author.")
}
```

`r fig_nums("fig_whc")`


# Empirical soil moisture stress functions

The set of empirical soil moisture stress functions $\beta$ are distinguished by their sensitivity to declining soil moisture, determined by the maximum $\beta$ reduction at low soil moisture $\beta_0 \equiv \beta(\theta \rightarrow 0)$. Here, we use data of the apparent soil moisture-related reduction in light use efficiency (fLUE [@Stocker2018-uf]) and approximate these by $\beta$ functions ($\beta \approx$fLUE ). Specifically, we use the apparent sensitivity (fLUE values at low observed soil moisture, termed fLUE$_0$) to inform $\beta_0$. [@Stocker2018-uf] further found that the sensitivity is not uniform across sites investigated therein, but exhibits a linear relationship with mean aridity $\alpha'$. Hence, we model the sensitivity $\beta_0$ as described also in the main text (Eq. 3):
$$
\beta_0 = p_0 + p_1 \alpha'
$$
Different approaches to determining $p_0$ and $p_1$ are implemented and lead to $\beta_a$, $\beta_b$, and $\beta_c$. These are described as follows. In all approaches, the mean aridity $\alpha'$ is calculated as the annual mean ratio of AET/PET, calculated from daily AET and PET values from simulations with the SPLASH model [@Davis2017-jo] (the same set of simulations as described in the Methods in the main text, section 'Global P-model simulations'). 

## Empirical soil moisture stress function $\beta_a$

Here, fLUE$_0$ is taken as the mean fLUE value of data in the lowest soil moisture bin ($0 < \theta\leq 0.25$) for each site. The set of sites used here corresponds to Group 1 in Table S1. Parameters $p_1$ and $p_2$ are calculated from a linear regression between each site's $\beta_0$ and $\alpha'$. This linear regression is illustrated by Fig. `r fig_nums("regr_aI", display="cite")`.

The centre of the lowest soil moisture bin defines $\theta_0=0.125$. $\theta^{\ast}$ was taken here as 0.75.


```{r linearfit_I, message=FALSE, warning=FALSE, echo=FALSE}
library(dplyr)
library(readr)
load( "data/data_aligned_agg.Rdata" )     # loads 'df_dday_agg', etc.
ddf <- df_dday_agg
rm("df_dday_agg")  # rename to be a bit more concise

## Use only sites where NN method worked (i.e. that had clear and identifiable soil moisture limitation)
successcodes <- read.csv( paste0( myhome, "/sofun/utils_sofun/analysis_sofun/fluxnet2015/successcodes.csv" ), as.is = TRUE )
do.sites <- dplyr::filter( successcodes, successcode==1 | successcode==2 )$mysitename
ddf <- ddf %>% filter( mysitename %in% do.sites )

## Merge mean annual alpha (AET/PET) values into this dataframe
load( "../sofun/utils_sofun/analysis_sofun/fluxnet2015/data/alpha_fluxnet2015.Rdata" )  # loads 'df_alpha'
ddf <- ddf %>% left_join( rename( df_alpha, meanalpha=alpha ), by="mysitename" )

## Merge vegetation class
siteinfo <- read_csv("../sofun/input_fluxnet2015_sofun/siteinfo_fluxnet2015_sofun.csv")
ddf <- ddf %>% select( -starts_with("classid")) %>% left_join( select( siteinfo, mysitename, classid ), by = "mysitename" )

source("get_linearfit.R")
source("plot_linearfit.R")

linearfit_I  <- get_linearfit( filter( ddf, ratio_obs_mod_pmodel < 5 ), monthly=FALSE )
save( linearfit_I,  file="data/linearfit_I.Rdata" )
plot_linearfit( linearfit_I )
fig_regr_aI <- fig_nums("regr_aI", caption=" Maximum apparent light use efficiency reduction at low soil moisture (fLUE$_0$) versus the ratio of annual mean actual over potential evapotranspiration (AET/PET). Points represent sites of Group 1 (see Table S1). The linear regression model is given by the equation and the line.")
```

`r fig_nums("regr_aI")`

In summary, this method derives the following fitting parameters $p_0 =$ `r format( coef(linearfit_I$linmod)[1], digits = 3 )` and $p_1 =$ `r format( coef(linearfit_I$linmod)[2], digits = 3 )`.

Given the estimate of $\beta_0$ as a function of $\alpha'$ at each site, we can calculate the soil moisture stress function using daily varying soil moisture data and evaluate whether accounting for this resolves the bias of modelled GPP at low soil moisture. In `r fig_nums("boxpl_aI", display="cite")`, we're assessing this with GPP predictions from the P-model.


```{r correct_linearfit_Ia, message=FALSE, warning=FALSE, echo=FALSE}
source("calc_flue_est_alpha.R")

## add estimated fLUE to ddf
ddf <- ddf %>% filter( mysitename!="US-Var") %>%  mutate( flue_est_I = calc_flue_est_alpha( soilm_mean, alpha=meanalpha, apar=coef(linearfit_I$linmod)[1], bpar=coef(linearfit_I$linmod)[2], cpar=0.125, dpar=0.75 ) ) %>%
                          mutate( bias_pmodel_diff = gpp_pmodel - gpp_obs,
                                  bias_pmodel_diff_corr = gpp_pmodel * fvar - gpp_obs,
                                  bias_pmodel_diff_corr_I = gpp_pmodel * flue_est_I - gpp_obs
                                 )

## evaluate bias~soil moisture
nbins <- 5
binwidth <- 1/nbins
bins <- seq( from=0, to=1, by=binwidth )
ddf <- ddf %>% mutate( inbin = cut( as.numeric(fvar), breaks = bins ) ) %>%
               mutate( lobin = ifelse( inbin %in% c("(0,0.2]", "(0.2,0.4]", "(0.4,0.6]", "(0.6,0.8]"), TRUE, FALSE ) )

## Bias (difference), correction by fLUE directly, fLUE bins
bp1 <- boxplot( bias_pmodel_diff ~ inbin, data=ddf, col="tomato", las=1, outline = FALSE, na.rm=TRUE, add=FALSE, staplewex=0, whisklty=1, at=(bins[1:nbins]+1/nbins*0.1), boxwex=0.03, xlim=c(0,1), xlab="fLUE bins", ylab=expression( paste("bias (gC m"^-2, "d"^-1, ")" ) ), staplewex=0, whisklty=1, ylim=c(-6,6) )
bp2 <- boxplot( bias_pmodel_diff_corr_I ~ inbin, data=ddf, col="springgreen3", las=1, outline = FALSE, na.rm=TRUE, add=TRUE, staplewex=0, whisklty=1, at=(bins[1:nbins]+1/nbins*0.3), boxwex=0.03, xlim=c(0,1), axes=FALSE, staplewex=0, whisklty=1, ylim=c(-6,6) )
bp3 <- boxplot( bias_pmodel_diff_corr ~ inbin, data=ddf, col="springgreen4", las=1, outline = FALSE, na.rm=TRUE, add=TRUE, staplewex=0, whisklty=1, at=(bins[1:nbins]+1/nbins*0.5), boxwex=0.03, xlim=c(0,1), axes=FALSE, staplewex=0, whisklty=1, ylim=c(-6,6) )
abline( h=0.0, lty=3 )
legend("bottomleft", c("P-model", expression("P-model, corrected by " ~ beta[a] ), "P-model, corrected by fLUE"), bty="n", fill=c("tomato", "springgreen3", "springgreen4") )
boxpl_aI <- fig_nums("boxpl_aI", caption=" Bias of GPP simulated by the P-model for different fLUE bins. Original P-model outputs are in red, outputs corrected by fLUE (stress multiplier taken as fLUE) in dark green, and outputs corrected by soil moisture stress function a in light green.")
```

`r fig_nums("boxpl_aI")`

<!-- Here are the statistics for I, mean across the lower five fLUE bins: -->

```{r echo=FALSE, message=FALSE, warning=FALSE}
## get statistics
suppressMessages(library(hydroGOF))
bybin_orig <- ddf %>% group_by( inbin ) %>% summarise( bias = mean(bias_pmodel_diff, na.rm=TRUE), rmse = rmse( bias_pmodel_diff + gpp_obs, gpp_obs) )
bybin_corr <- ddf %>% group_by( inbin ) %>% summarise( bias = mean(bias_pmodel_diff_corr, na.rm=TRUE), rmse = rmse( bias_pmodel_diff_corr + gpp_obs, gpp_obs) )
bybin_corI <- ddf %>% group_by( inbin ) %>% summarise( bias = mean(bias_pmodel_diff_corr_I, na.rm=TRUE), rmse = rmse( bias_pmodel_diff_corr_I + gpp_obs, gpp_obs) )

# print(paste("With fLUE, mean bias is reduced from", format(mean(bybin_orig$bias[1:5]), digits=3), "to", format(mean(bybin_corr$bias[1:5]), digits=3)))
# print(paste("With fLUE, RMSE is reduced from", format(mean(bybin_orig$rmse[1:5]), digits=3), "to", format(mean(bybin_corr$rmse[1:5]), digits=3)))
#
# print(paste("With I, mean bias is reduced from", format(mean(bybin_orig$bias[1:5]), digits=3), "to", format(mean(bybin_corI$bias[1:5]), digits=3)))
# print(paste("With I, RMSE is reduced from", format(mean(bybin_orig$rmse[1:5]), digits=3), "to", format(mean(bybin_corI$rmse[1:5]), digits=3)))

## save performance statistics
df_corI <- ddf %>% summarise( bias = mean(bias_pmodel_diff_corr_I, na.rm=TRUE), rmse = rmse( bias_pmodel_diff_corr_I + gpp_obs, gpp_obs) )
df_orig <- ddf %>% summarise( bias = mean(bias_pmodel_diff, na.rm=TRUE), rmse = rmse( bias_pmodel_diff + gpp_obs, gpp_obs) )
df_coef   <-  tibble( approach="original",
                      apar=NA,
                      bpar=NA,
                      apar_grass=NA, bpar_grass=NA,
                      bias=df_orig$bias, rmse=df_orig$rmse
                      ) %>%
              bind_rows(
                tibble( approach="a",
                        apar=coef(linearfit_I$linmod)["(Intercept)"],
                        bpar=coef(linearfit_I$linmod)["meanalpha"],
                        apar_grass=NA, bpar_grass=NA,
                        bias=df_corI$bias, rmse=df_corI$rmse
                        ))
```

`r fig_nums("boxpl_aI", display="cite")` shows that the bias in predicted GPP is reduced but not fully resolved. Hence, we define alternative, more sensitive stress functions below.

## Empirical soil moisture stress function $\beta_b^{\ast}$

As for $\beta_a$, approach for fitting $\beta_b^{\ast}$ is also based on the relationship between minimum fLUE (fLUE$_0$) and mean AET/PET. However, in contrast to $\beta_a$, fLUE$_0$ is derived here from fitting a quadratic function of the form described above to fLUE data for each site individually and taking the y-axis intersect of the site-specific fitted function to define fLUE$_0$ ($=\beta_0$). Then, this $\beta_0$ is regressed against $\alpha'$ as above.

Here, $\theta_0=0.0$ and $\theta^{\ast}=0.9$.

```{r get_linearfit_II, warning=FALSE, message=FALSE, echo=FALSE}
## load additional functions
source("get_linearfit_II.R")
source("stress_quad_1sided.R")

## Use fixed point (soil moisture) above which stress function is 1 (no stress)
x0_fix <- 0.9
linearfit_II <- get_linearfit_II( filter( ddf, dday>=0 ), target="fvar", monthly=FALSE, bin=FALSE, x0_fix=x0_fix, useweights = TRUE, doplot = FALSE )
save( linearfit_II, file="data/linearfit_II_ratio.Rdata" )
```


```{r correct_linearfit_II_boxpl, warning=FALSE, message=FALSE, echo=FALSE}
## add estimated fLUE to ddf
ddf <- ddf %>%  mutate( flue_est_II = stress_quad_1sided_alpha( soilm_mean, meanalpha, x0_fix, coef(linearfit_II$linmod)[["(Intercept)"]], coef(linearfit_II$linmod)[["meanalpha"]] ) ) %>%
                mutate( bias_pmodel_diff_corr_II = gpp_pmodel * flue_est_II - gpp_obs )

## Bias (difference), correction by fLUE directly, fLUE bins -->
bp1 <- boxplot( bias_pmodel_diff ~ inbin, data=ddf, col="tomato", las=1, outline = FALSE, na.rm=TRUE, add=FALSE, staplewex=0, whisklty=1, at=(bins[1:nbins]+1/nbins*0.1), boxwex=0.03, xlim=c(0,1), xlab="fLUE bins", ylab=expression( paste("bias (gC m"^-2, "d"^-1, ")" ) ), staplewex=0, whisklty=1, ylim=c(-6,6) )
bp2 <- boxplot( bias_pmodel_diff_corr_II ~ inbin, data=ddf, col="springgreen3", las=1, outline = FALSE, na.rm=TRUE, add=TRUE, staplewex=0, whisklty=1, at=(bins[1:nbins]+1/nbins*0.3), boxwex=0.03, xlim=c(0,1), axes=FALSE, staplewex=0, whisklty=1, ylim=c(-6,6) )
bp3 <- boxplot( bias_pmodel_diff_corr ~ inbin, data=ddf, col="springgreen4", las=1, outline = FALSE, na.rm=TRUE, add=TRUE, staplewex=0, whisklty=1, at=(bins[1:nbins]+1/nbins*0.5), boxwex=0.03, xlim=c(0,1), axes=FALSE, staplewex=0, whisklty=1, ylim=c(-6,6) )
abline( h=0.0, lty=3 )
legend("bottomleft", c("P-model", expression("P-model, corrected by " ~ beta[b] ~ "*" ), "P-model, corrected by fLUE"), bty="n", fill=c("tomato", "springgreen3", "springgreen4") )
boxpl_aII <- fig_nums("boxpl_aII", caption=" Bias of GPP simulated by P-model for different fLUE bins. Original P-model outputs are in red, outputs corrected by fLUE (stress multiplier taken as fLUE) in dark green, and outputs corrected by soil moisture stress function b* in light green.")
```

`r fig_nums("boxpl_aII")`

This approach is more effective in reducing the bias in the lower four fLUE bins than $\beta_a$, but corrected values still have the tendency to be biased high in the lowest fLUE bin and it also introduces a negative bias in the highest fLUE bin.

The following fitting parameters are derived: $p_0 =$ `r format( coef(linearfit_II$linmod)[1], digits = 3 )` and $p_1 =$ `r format( coef(linearfit_II$linmod)[2], digits = 3 )`.

**This method ($\beta_b^{\ast}$) is not used for further analysis.** Instead, an alternative intermediate soil moisture stress function $\beta_b$ is used (described below).

<!-- Here are the statistics for $\beta_b$, mean across the lower five fLUE bins: -->

```{r echo=FALSE}
## get statistics
bybin_orig <- ddf %>% group_by( inbin ) %>% summarise( bias = mean(bias_pmodel_diff, na.rm=TRUE), rmse = rmse( bias_pmodel_diff + gpp_obs, gpp_obs) )
bybin_corr <- ddf %>% group_by( inbin ) %>% summarise( bias = mean(bias_pmodel_diff_corr, na.rm=TRUE), rmse = rmse( bias_pmodel_diff_corr + gpp_obs, gpp_obs) )
bybin_corII<- ddf %>% group_by( inbin ) %>% summarise( bias = mean(bias_pmodel_diff_corr_II, na.rm=TRUE), rmse = rmse( bias_pmodel_diff_corr_II + gpp_obs, gpp_obs) )

# print(paste("With fLUE, mean bias is reduced from", format(mean(bybin_orig$bias[1:5]), digits=3), "to", format(mean(bybin_corr$bias[1:5]), digits=3)))
# print(paste("With fLUE, RMSE is reduced from", format(mean(bybin_orig$rmse[1:5]), digits=3), "to", format(mean(bybin_corr$rmse[1:5]), digits=3)))
# 
# print(paste("With II, mean bias is reduced from", format(mean(bybin_orig$bias[1:5]), digits=3), "to", format(mean(bybin_corII$bias[1:5]), digits=3)))
# print(paste("With II, RMSE is reduced from", format(mean(bybin_orig$rmse[1:5]), digits=3), "to", format(mean(bybin_corII$rmse[1:5]), digits=3)))

## save performance statistics
df_corII <- ddf %>% summarise( bias = mean(bias_pmodel_diff_corr_II, na.rm=TRUE), rmse = rmse( bias_pmodel_diff_corr_II + gpp_obs, gpp_obs) )
df_coef <- df_coef %>% bind_rows( tibble(
  approach="b*",
  apar=coef(linearfit_II$linmod)["(Intercept)"],
  bpar=coef(linearfit_II$linmod)["meanalpha"],
  apar_grass=NA, bpar_grass=NA,
  bias=df_corII$bias, rmse=df_corII$rmse
))
```

## Empirical soil moisture stress function $\beta_c$

In view of the remaining positive bias in the lowest fLUE bin, we derive soil moisture stress functions with an even stronger sensitivity. To achieve that, we use only data from sites where approach $\beta_b^{\ast}$ does not effectively remove the bias. The method for finding fit parameters itself is identical to approach taken for $\beta_b^{\ast}$ (also with $\theta_0=0.0$ and $\theta^{\ast}=0.9$). Sites used here are IT-Noe, FR-Pue, AU-Stp, AU-Fog, AU-DaP, AU-ASM, IT-SRo, US-SRG, US-SRM, and US-Var.

```{r get_linearfit_III, warning=FALSE, message=FALSE, echo=FALSE}
## Use fixed point (soil moisture) above which stress function is 1 (no stress)
x0_fix <- 0.9
notorious <- c("IT-Noe", "FR-Pue", "AU-Stp", "AU-Fog", "AU-DaP", "AU-ASM", "IT-SRo", "US-SRG", "US-SRM", "US-Var")
linearfit_III <- get_linearfit_II( filter( ddf, dday>=0, mysitename %in% notorious ), target="fvar", monthly=FALSE, bin=FALSE, x0_fix=x0_fix, useweights = TRUE, doplot = FALSE )
save( linearfit_III, file="data/linearfit_III.Rdata" )
```

```{r correct_linearfit_III_boxpl, warning=FALSE, message=FALSE, echo=FALSE}
## add estimated fLUE to ddf
ddf <- ddf %>%  mutate( flue_est_III = stress_quad_1sided_alpha( soilm_mean, meanalpha, x0_fix, coef(linearfit_III$linmod)[["(Intercept)"]], coef(linearfit_III$linmod)[["meanalpha"]] ) ) %>%
                mutate( bias_pmodel_diff_corr_III = gpp_pmodel * flue_est_III - gpp_obs )

## Bias (difference), correction by fLUE directly, fLUE bins
bp1 <- boxplot( bias_pmodel_diff ~ inbin, data=ddf, col="tomato", las=1, outline = FALSE, na.rm=TRUE, add=FALSE, staplewex=0, whisklty=1, at=(bins[1:nbins]+1/nbins*0.1), boxwex=0.03, xlim=c(0,1), xlab="fLUE bins", ylab=expression( paste("bias (gC m"^-2, "d"^-1, ")" ) ), staplewex=0, whisklty=1, ylim=c(-6,6) )
bp2 <- boxplot( bias_pmodel_diff_corr_III ~ inbin, data=ddf, col="springgreen3", las=1, outline = FALSE, na.rm=TRUE, add=TRUE, staplewex=0, whisklty=1, at=(bins[1:nbins]+1/nbins*0.3), boxwex=0.03, xlim=c(0,1), axes=FALSE, staplewex=0, whisklty=1, ylim=c(-6,6) )
bp3 <- boxplot( bias_pmodel_diff_corr ~ inbin, data=ddf, col="springgreen4", las=1, outline = FALSE, na.rm=TRUE, add=TRUE, staplewex=0, whisklty=1, at=(bins[1:nbins]+1/nbins*0.5), boxwex=0.03, xlim=c(0,1), axes=FALSE, staplewex=0, whisklty=1, ylim=c(-6,6) )
abline( h=0.0, lty=3 )
legend("bottomleft", c("P-model", expression("P-model, corrected by " ~ beta[c] ), "P-model, corrected by fLUE"), bty="n", fill=c("tomato", "springgreen3", "springgreen4") )
boxpl_aIII <- fig_nums("boxpl_aIII", caption=" Bias of GPP simulated by P-model for different fLUE bins. Original P-model outputs are in red, outputs corrected by fLUE (stress multiplier taken as fLUE) in dark green, and outputs corrected by soil moisture stress function c in light green.")
```

`r fig_nums("boxpl_aIII")`

<!-- Here are the statistics for $\beta_c$, mean across the lower five fLUE bins: -->

```{r echo=FALSE}
## get statistics
bybin_orig  <- ddf %>% group_by( inbin ) %>% summarise( bias = mean(bias_pmodel_diff, na.rm=TRUE), rmse = rmse( bias_pmodel_diff + gpp_obs, gpp_obs) )
bybin_corr  <- ddf %>% group_by( inbin ) %>% summarise( bias = mean(bias_pmodel_diff_corr, na.rm=TRUE), rmse = rmse( bias_pmodel_diff_corr + gpp_obs, gpp_obs) )
bybin_corIII<- ddf %>% group_by( inbin ) %>% summarise( bias = mean(bias_pmodel_diff_corr_III, na.rm=TRUE), rmse = rmse( bias_pmodel_diff_corr_III + gpp_obs, gpp_obs) )

# print(paste("With fLUE, mean bias is reduced from", format(mean(bybin_orig$bias[1:5]), digits=3), "to", format(mean(bybin_corr$bias[1:5]), digits=3)))
# print(paste("With fLUE, RMSE is reduced from",      format(mean(bybin_orig$rmse[1:5]), digits=3), "to", format(mean(bybin_corr$rmse[1:5]), digits=3)))
# 
# print(paste("With III, mean bias is reduced from", format(mean(bybin_orig$bias[1:5]), digits=3), "to", format(mean(bybin_corIII$bias[1:5]), digits=3)))
# print(paste("With III, RMSE is reduced from",      format(mean(bybin_orig$rmse[1:5]), digits=3), "to", format(mean(bybin_corIII$rmse[1:5]), digits=3)))

## save performance statistics
df_corIII <- ddf %>% summarise( bias = mean(bias_pmodel_diff_corr_III, na.rm=TRUE), rmse = rmse( bias_pmodel_diff_corr_III + gpp_obs, gpp_obs) )
df_coef <- df_coef %>% bind_rows( tibble(
  approach="c",
  apar=coef(linearfit_III$linmod)["(Intercept)"],
  bpar=coef(linearfit_III$linmod)["meanalpha"],
  apar_grass=NA, bpar_grass=NA,
  bias=df_corIII$bias, rmse=df_corIII$rmse
))
```

## Empirical soil moisture stress function $\beta_b$

In view of the trade-off between resolving the bias in the lowest and highest fLUE bins and remaining biases, we developed a method that uses additional information to distinguish between vegetation types for determining the sensitivity of the soil moisture stress function. We applied the same method as in approach II, but determined fit parameters for grasslands (IGBP vegetation type GRA or CSH) and other ecosystems (other vegetation types) separately. As for methods $\beta_a$ and $\beta_c$, $\theta_0=0.0$ and $\theta^{\ast}=0.9$.

```{r get_linearfit_IV, warning=FALSE, message=FALSE, echo=FALSE}
## Merge IGBP class ID into dataframe
siteinfo <- read_csv("../sofun/input_fluxnet2015_sofun/siteinfo_fluxnet2015_sofun.csv")
ddf <- ddf %>% select( -starts_with("classid")) %>% left_join( select( siteinfo, mysitename, classid ), by = "mysitename" )

## load additional functions
source("get_linearfit_IV.R")
source("stress_quad_1sided.R")

## Use fixed point (soil moisture) above which stress function is 1 (no stress)
x0_fix <- 0.9
linearfit_IV <- get_linearfit_IV( filter( ddf, dday >=0 & !(mysitename %in% c("CH-Oe1", "US-Var")) ), target="fvar", monthly=FALSE, bin=FALSE, x0_fix=x0_fix, useweights = TRUE, doplot = FALSE )
save( linearfit_IV, file="data/linearfit_IV.Rdata" )
```

Again, we have a linear relationship between the maximum LUE reduction (y-axis intersect) and mean site aridity (AET/PET), as shown in `r fig_nums("regr_aIV", display="cite")`.

```{r echo=FALSE}
par(las=1)
with( filter( linearfit_IV$data, classid %in% c("GRA", "CSH")), plot( meanalpha, y0, pch=16 , ylim=c(0,1), xlim=c(0,1), xlab="AET/PET", ylab=expression(paste("fLUE"[0]))) )
with( filter( linearfit_IV$data, !(classid %in% c("GRA", "CSH"))), points( meanalpha, y0, pch=1 ) )
abline( linearfit_IV$linmod_grass )
abline( linearfit_IV$linmod_tree, lty=2 )
legend("topleft", c("Grasslands and open shrublands (GRA, CSH)", "Other ecosystems"), pch=c(16, 1), bty="n")
regr_aIV <- fig_nums("regr_aIV", caption=" Maximum apparent light use efficiency reduction at low soil moisture (fLUE$_0$) versus the ratio of annual mean actual over potential evapotranspiration (AET/PET). Points represent sites of group 1 (see Table S1). Note that values of fLUE$_0$ are not identical to the values used in Fig. S13 due to different definitions.")
```

`r fig_nums("regr_aIV")`

This leads to an effective bias reduction, as shown in `r fig_nums( "boxpl_aIV", display = "cite")`.

```{r correct_linearfit_IV, warning=FALSE, message=FALSE, echo=FALSE}
source("stress_quad_1sided.R")
apar <- c( coef(linearfit_IV$linmod_tree)[["(Intercept)"]], coef(linearfit_IV$linmod_grass)[["(Intercept)"]] )
bpar <- c( coef(linearfit_IV$linmod_tree)[["meanalpha"]], coef(linearfit_IV$linmod_grass)[["meanalpha"]] )

## add estimated fLUE to ddf
ddf$flue_est_IV <- rep( NA, nrow( ddf ))
for (idx in seq(nrow(ddf))){
  ddf$flue_est_IV[idx] <- stress_quad_1sided_alpha_grasstree( ddf$soilm_mean[idx], ddf$meanalpha[idx], x0 = 0.9, apar, bpar, ddf$classid[idx] )
}
ddf <- ddf %>% mutate( bias_pmodel_diff_corr_IV = gpp_pmodel * flue_est_IV - gpp_obs )

## Bias (difference), correction by fLUE directly, fLUE bins
bp1 <- boxplot( bias_pmodel_diff ~ inbin, data=ddf, col="tomato", las=1, outline = FALSE, na.rm=TRUE, add=FALSE, staplewex=0, whisklty=1, at=(bins[1:nbins]+1/nbins*0.1), boxwex=0.03, xlim=c(0,1), xlab="fLUE bins", ylab=expression( paste("bias (gC m"^-2, "d"^-1, ")" ) ), staplewex=0, whisklty=1, ylim=c(-6,6) )
bp2 <- boxplot( bias_pmodel_diff_corr_IV ~ inbin, data=ddf, col="springgreen3", las=1, outline = FALSE, na.rm=TRUE, add=TRUE, staplewex=0, whisklty=1, at=(bins[1:nbins]+1/nbins*0.3), boxwex=0.03, xlim=c(0,1), axes=FALSE, staplewex=0, whisklty=1, ylim=c(-6,6) )
bp3 <- boxplot( bias_pmodel_diff_corr ~ inbin, data=ddf, col="springgreen4", las=1, outline = FALSE, na.rm=TRUE, add=TRUE, staplewex=0, whisklty=1, at=(bins[1:nbins]+1/nbins*0.5), boxwex=0.03, xlim=c(0,1), axes=FALSE, staplewex=0, whisklty=1, ylim=c(-6,6) )
abline( h=0.0, lty=3 )
legend("bottomleft", c("P-model", expression("P-model, corrected by " ~ beta[b] ), "P-model, corrected by fLUE"), bty="n", fill=c("tomato", "springgreen3", "springgreen4") )
boxpl_aIV <- fig_nums("boxpl_aIV", caption=" Bias of GPP simulated by P-model for different fLUE bins. Original P-model outputs are in red, outputs corrected by fLUE (stress multiplier taken as fLUE) in dark green, and outputs corrected by soil moisture stress function b in light green.")
```

`r fig_nums("boxpl_aIV")`

<!-- Here are the statistics for IV, mean across the lower five fLUE bins: -->

```{r echo=FALSE}
## get statistics
bybin_orig  <- ddf %>% group_by( inbin ) %>% summarise( bias = mean(bias_pmodel_diff, na.rm=TRUE), rmse = rmse( bias_pmodel_diff + gpp_obs, gpp_obs) )
bybin_corr  <- ddf %>% group_by( inbin ) %>% summarise( bias = mean(bias_pmodel_diff_corr, na.rm=TRUE), rmse = rmse( bias_pmodel_diff_corr + gpp_obs, gpp_obs) )
bybin_corIV <- ddf %>% group_by( inbin ) %>% summarise( bias = mean(bias_pmodel_diff_corr_IV , na.rm=TRUE), rmse = rmse( bias_pmodel_diff_corr_IV  + gpp_obs, gpp_obs) )

# print(paste("With fLUE, mean bias is reduced from", format(mean(bybin_orig$bias[1:5]), digits=3), "to", format(mean(bybin_corr$bias[1:5]), digits=3)))
# print(paste("With fLUE, RMSE is reduced from",      format(mean(bybin_orig$rmse[1:5]), digits=3), "to", format(mean(bybin_corr$rmse[1:5]), digits=3)))
#
# print(paste("With IV , mean bias is reduced from", format(mean(bybin_orig$bias[1:5]), digits=3), "to", format(mean(bybin_corIV $bias[1:5]), digits=3)))
# print(paste("With IV , RMSE is reduced from",      format(mean(bybin_orig$rmse[1:5]), digits=3), "to", format(mean(bybin_corIV $rmse[1:5]), digits=3)))

## save performance statistics
df_corIV <- ddf %>% summarise( bias = mean(bias_pmodel_diff_corr_IV, na.rm=TRUE), rmse = rmse( bias_pmodel_diff_corr_IV + gpp_obs, gpp_obs) )
df_coef <- df_coef %>% bind_rows( tibble(
  approach="b",
  apar=coef(linearfit_IV$linmod_tree)["(Intercept)"],
  bpar=coef(linearfit_IV$linmod_tree)["meanalpha"],
  apar_grass=coef(linearfit_IV$linmod_grass)["(Intercept)"],
  bpar_grass=coef(linearfit_IV$linmod_grass)["meanalpha"],
  bias=df_corIV$bias, rmse=df_corIV$rmse
 ))
```


## Summary of approaches

For further analyses, we only retained $\beta_a$, $\beta_b$, and $\beta_c$ (referred to as Approaches a, b, and c in Table S2.). Approach b yields the best performance statistics in removing the bias in simulated GPP (P-model outputs assessed here) and serves as a best estimate. Approaches a and c provide a lower and upper boundary for the uncertainty in the sensitivity of GPP (or LUE) to declining soil moisture. An overview of fitted parameters $p_0$ and $p_1$ (p_0 and p_1 in Table S2) is given below. Note that approaches a, b, and c are used for global P-model simulations s1a, s1b, and s1c, respectively.
```{r echo = FALSE, results = 'asis', warning=FALSE}
df_coef <- df_coef %>% mutate( Simulation=c(NA, "s1a", NA, "s1c", "s1b") )
df_coef %>% mutate( apar=format(apar, digits=1), bpar=format(bpar, digits=3),
                    apar_grass=format(apar_grass, digits=3), bpar_grass=format(bpar_grass, digits=2)) %>%
            rename( a=apar, b=bpar, a_grass=apar_grass, b_grass=bpar_grass) %>%
            select( Approach=approach, Simulation, p_0=a, p_1=b, p_0_grass=a_grass, p_1_grass=b_grass, Bias=bias, RMSE=rmse ) %>%
            knitr::kable( caption = " Overview of parameters for the empirical soil moisture stress functions and performance metrics of corrected GPP with respective soil moisture stress functions following approaches a, b, and c.")
## save stuff
save( ddf, file = "data/ddf.Rdata" )
save( df_coef, file = "data/df_coef.Rdata")
```


# References



